### 2015-02-28 curriculo/cv.tex
[ADD] added cv

%!TEX TS-program = xelatex
\documentclass[]{friggeri-cv}
\addbibresource{bibliography.bib}

\begin{document}
\header{Henrique}{Rocha}
       {Telecommunications and Informatics}

% In the aside, each new line forces a line break
\begin{aside}
  \section{about}
    Henrique Duarte Lopes Rocha
    ~
    Rua dos Cedros nº100
    Alto do Lagoal
    2760-025 Caxias
    Portugal
    ~
    (+351) 938781922
    ~
    \href{mailto:hdlopesrocha91@gmail.com}{hdlopesrocha91\\@gmail.com}
    \href{http://web.ist.utl.pt/ist168621}{WebSite}
    \href{http://www.youtube.com/user/hdlopesrocha}{YouTube}
  \section{languages}
    portuguese (native)
    english (fluent)
  \section{programming}
    {\color{red} $\varheartsuit$} Java
	 (Android, PlayFramework, LWJGL, MapReduce), C, C++ (OpenGL), C\# (XNA, Windows Forms, WPF, lpsolve), VB, Python, Scala, Scheme, HTML, CSS, JavaScript (jQuery, AJAX, WebGL), SQL, PHP, GLSL, HLSL.  
   \section{databases}
	MySQL, PostgreSQL, MongoDB, DynamoDB.
   \section{soft skills}
	Time Management
	Problem-Solving
	Team Player
	Self Confidence
	Adaptability
\end{aside}

\section{education}

\begin{entrylist}
   \entry
    {2013–now}
    {Master of Science}
    {Instituto Superior Técnico - Taguspark}
    {Telecommunications and Informatics Engineering}

   \entry
    {2009–2014}
    {Bachelor of Science}
    {Instituto Superior Técnico - Taguspark}
    {Communication Networks Engineering}

  \entry
    {2006–2009}
    {Secondary School}
    {Escola Secundária Luís de Freitas Branco, Paço de Arcos}
    {Science \& Technology}

\end{entrylist}

\section{experience}

\begin{entrylist}
  \entry
    {11 2014\\(ongoing)}
    {Web Development}
    {Bullray}
    {\emph{Development of an architecture for a chat application. Experience with Android, PlayFramework and MongoDB.}}
  \entry
    {04 2014\\(6 months)}
    {Scientific Initiation Scholarship}
    {INESC}
    {\emph{Frontend Development for Provide Results's Bimk project. Experience with PlayFramework, Revit API and GIT.}}
  \entry
    {09 2013\\(6 months)}
    {Scientific Initiation Scholarship}
    {CEG-IST}
    {\emph{Programming Decision Analysis Software (MACBETH). Development of an algorithm to suggest judgements based on past decisions using \emph{lpsolve}.}}
  \entry
    {08 2013\\(2 weeks)}
    {Summer Work}
    {LCG}
    {\emph{Frontend development for Decision Analysis Software using \emph{ASP.NET}. }}
  \entry
    {11 2012}
    {Pizza Night Competition}
    {Microsoft Lisbon Experience}
    {\emph{2º place on Windows Phone Category} with \emph{Infinity Edge} game.}
  \entry
    {08 2006\\(1 month)}
    {Garbage Collector}
    {Jovens em Movimento, Oeiras}
    {Cleaning and maintenance of public spaces.}

\end{entrylist}

\section{applications}

\begin{entrylist}
  \entry
    {2013}
    {OpenGlobe}
    {\href{http://web.ist.utl.pt/ist168621/?page=10}{Instituto Superior Técnico}}
    {Modeling Earth surface with GPS devices and topographic collected data. It is also possible to predict satellites position through ephemeris data.}

  \entry
    {2012}
    {Infinity Edge (Windows Phone 7)}
    {\href{http://www.windowsphone.com/en-us/store/app/infinity-edge/05516463-3a89-4351-996e-62e5b4519aeb}{windows marketplace}}
    {First Person Shooter with spaceships.}

  \entry
    {2011}
    {Era of Empires}
    {\href{http://web.ist.utl.pt/ist168621/?page=8}{Personal Project}}
    {3D environment generator and Real Time Strategy game engine.}

\end{entrylist}

\section{interests}
    3D Studio Max, Photoshop, Game Development, Web Design, Network Security, Jogging and Yoga.

%\section{publications}

%Put your publications here!

%%% This piece of code has been commented by Karol Kozioł due to biblatex errors. 
% 
%\printbibsection{article}{article in peer-reviewed journal}
%\begin{refsection}
%  \nocite{*}
%  \printbibliography[sorting=chronological, type=inproceedings, title={international peer-reviewed conferences/proceedings}, notkeyword={france}, heading=subbibliography]
%\end{refsection}
%\begin{refsection}
%  \nocite{*}
%  \printbibliography[sorting=chronological, type=inproceedings, title={local peer-reviewed conferences/proceedings}, keyword={france}, heading=subbibliography]
%\end{refsection}
%\printbibsection{misc}{other publications}
%\printbibsection{report}{research reports}

\end{document}