### 2015-03-05 tese/state.tex
[ADD] added first tex (inro on NAT), added a paper about IPv4 (references must be done carefully)

\documentclass{article}

\usepackage[brazilian]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{color}
\usepackage{makeidx}
\usepackage{float}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[hmargin=3cm,vmargin=3.5cm]{geometry}
\usepackage{indentfirst}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\usepackage{eurosym}
\usepackage{url}

\lstset{frame=tb,
  language=C,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true
  tabsize=3
}

\usepackage{titlesec}
\titleformat{\section}{\hrule \large\bfseries}{\thesection}{1em}{}

\title{Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração potenciado pela Tecnologia WebRTC}
\author{
	Henrique Lopes Rocha \\ hdlopesrocha91@gmail.com
}

\date{Fevereiro 2015}
\begin{document}
	\maketitle
	\newpage
	\tableofcontents
	\newpage

\textbf{Abstract}
\section{Introduction}
The need to build a global comunication network in an era when, almost nobody had access to it, caused that some protocols weren't suitable for a huge increase on the amount of publicly know users. IPv4 limits the number o public addresses in such a way that today are scarse. One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group though their private addresses, this technique is also know as Network Address Translation (NAT).

\begin{thebibliography}{9}

\bibitem{activator} Typesafe - Build Reactive Applications on the JVM - \url{http://typesafe.com/}
\bibitem{github} GitHub - Build software better, together - \url{http://github.com/}
\bibitem{gitlab} GitLab - Version Control on your Server - \url{http://gitlab.com/}
\bibitem{gitolite} Gitolite - Hosting Git Repositories - \url{http://gitolite.com/}
\bibitem{gitweb} Gitweb - Git SCM Wiki - \url{http://git.wiki.kernel.org/index.php/Gitweb}
\bibitem{iptables} iptables - Linux man page - \url{http://linux.die.net/man/8/iptables}
\bibitem{jquery} jQuery - write less, do more - \url{http://jquery.com/}
\bibitem{json} JSON - JavaScript Object Notation - \url{http://json.org/}
\bibitem{mongodb} MongoDB - NoSQL database - \url{http://mongodb.org/}
\bibitem{rest} REST - Representational state transfer - \url{http://en.wikipedia.org/wiki/Representational_state_transfer}
\bibitem{sdk} Android SDK - Software Development Kit - \url{http://developer.android.com/sdk/index.html}
\bibitem{ubuntu} Ubuntu Server - for scale-out computing - \url{http://www.ubuntu.com/server}
\bibitem{virtualbox} VirtualBox - Oracle VM - \url{http://virtualbox.org/}

\end{thebibliography}

\end{document}