### 2015-05-27 tese/method.tex
need conclusions and abstract

\subsection{Scientific Methodology} % Section

The most important task in order to implement our project is the infrastructure, without it our project wouldn't work. The basic infrastructure will be our first concern, namely the Web Server configuration, ICE and signaling servers.

The Web development can only start after the basic infrastructure configuration. When leading with streams, the infrastructure will be complemented with stream servers.

There will be four prototypes, the first one should implement most of streaming and collaborative functionalities. The second prototype will have the same functionalities but a more friendly user interface which will be tested and reviewed by potential users. The third prototype will be an improvement based on the users feedback. In the final prototype we will improve our system performance.

Briefly, our web application will be tested with real users and unit tests, the system in general will be validated with benchmarks such as analyzing the its behavior and delays by growing the amount of users and data.