### 2015-05-29 tese/schd01.tex

\tikzset{every picture/.style={xscale=0.51,yscale=0.51,transform shape}}

\begin{ganttchart}[vgrid, hgrid]{1}{36}
\gantttitle{2015}{28}
\gantttitle{2016}{8} \\
\gantttitle{Jun}{4}
\gantttitle{Jul}{4}
\gantttitle{Aug}{4}
\gantttitle{Sep}{4}
\gantttitle{Oct}{4}
\gantttitle{Nov}{4}
\gantttitle{Dec}{4}
\gantttitle{Jan}{4}
\gantttitle{Feb}{4} \\
\gantttitlelist{1,...,36}{1}\\
%First Group
\ganttgroup{Thesis}{1}{32} \\
\ganttbar{Related work evaluation}{1}{28}\\
\ganttbar{Essay writing}{19}{32} \\
\ganttbar{Paper writing}{25}{32} \\
\ganttmilestone{Delivery}{35}{35} \\
%\ganttlink{elem0}{elem1}
%\ganttgroup{Resume}{1}{10} \\

%\ganttmilestone{Milestone 1}{11}
%Second Group
\ganttgroup{Implementation}{1}{1}\ganttgroup{}{4}{8}\ganttgroup{}{11}{26}\ganttgroup{}{29}{30} \\
\ganttbar{Basic configuration}{1}{1} \\
%\ganttlink{elem4}{elem5}
%\ganttmilestone{Milestone 1}{11}
%Third Group
\ganttbar{WebRTC basics}{4}{4} \\
\ganttbar{Video \& Audio Communication}{4}{5} \\
\ganttbar{Non-Interactive Record \& Playback}{6}{8} \\
\ganttbar{Interactive Record \& Playback}{11}{14}\\
\ganttmilestone{First Prototype}{18}\\
\ganttbar{Collaborative Environment}{15}{18}\\
\ganttmilestone{Second Prototype}{22} \\
\ganttbar{Interface Improvement I}{19}{22} \\
\ganttbar{Usability Tests I}{23}{23} \\
\ganttmilestone{Third Prototype}{25} \\
\ganttbar{Interface Improvement II}{24}{25} \\
\ganttbar{Usability Tests II}{26}{26} \\
\ganttbar{Performance Tests}{29}{30} \\
\ganttbar{Performance Improvements}{30}{30} \\
\ganttmilestone{Final Prototype}{30}

%\ganttlink{elem8}{elem9}

\ganttlink{elem9}{elem10}
\ganttlink[link type=F-S]{elem11}{elem12}
\ganttlink{elem12}{elem13}
\ganttlink[link type=F-S]{elem13}{elem15}
\ganttlink[link type=F-S]{elem15}{elem17}
\ganttlink[link type=F-S]{elem17}{elem18}
\ganttlink[link type=F-S]{elem18}{elem20}
\ganttlink[link type=F-S]{elem20}{elem21}
\end{ganttchart}