### 2015-06-02 tese/project.tex
Added a few sentences to the abstract.

\documentclass{llncs}
\RequirePackage[english]{babel} % Se fizerem o texto em ingles
\usepackage[english]{babel}   
%\usepackage[latin1]{inputenc}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

\usepackage[fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{subfig}  % allows for indexgeneration
\usepackage{color}  % allows for indexgeneration
\usepackage{url}
\usepackage{hyperref}
\usepackage{float}
\usepackage{mathtools}
\usepackage{acronym}
\usepackage{pgfgantt}
\usepackage{adjustbox}
\usepackage{rotating}
\usepackage{dirtytalk}
\usepackage[graphicx]{realboxes}
\definecolor{fade}{rgb}{0.8,0.8,0.8}

\newacro{AES}{Advanced Encryption Standard}
\newacro{API}{Application Programming Interface}
\newacro{BOSH}{Bidirectional-streams Over Synchronous HTTP}
\newacro{CBC}{Cipher Block Chaining}
\newacro{CPU}{Central Processing Unit}
\newacro{CRUD}{Create, Read, Update and Delete}
\newacro{CSS}{Cascading Style Sheets}
\newacro{DOM}{Document Object Model}
\newacro{DTLS}{Datagram Transport Layer Security}
\newacro{DVD}{Digital Versatile Discs}
\newacro{FMS}{Flash Media Server}
\newacro{HTML}{HyperText Markup Language}
\newacro{HTTP}{Hypertext Transfer Protocol}
\newacro{ICE}{Interactive Connectivity Establishment}
\newacro{IPv4}{Internet Protocol Version 4}
\newacro{IPv6}{Internet Protocol Version 6}
\newacro{IP}{Internet Protocol}
\newacro{JSON}{JavaScript Object Notation}
\newacro{MPEG}{Moving Picture Experts Group}
\newacro{MVC}{Model-View-Controller}
\newacro{NAT}{Network Address Translation}
\newacro{OT}{Operational transformation}
\newacro{P2P}{Peer-to-peer}
\newacro{PSTN}{Public Switched Telephone Network}
\newacro{RFC}{Request For Comments}
\newacro{RTCP}{Real Time Control Protocol}
\newacro{RTSP}{Real Time Streaming Protocol}
\newacro{RTMFP}{Real-Time Media Flow Protocol}
\newacro{RTMP}{Real Time Messaging Protocol}
\newacro{RTP}{Real-time Transport Protocol}
\newacro{SAMI}{Synchronized Accessible Media Interchange}
\newacro{SCTP}{Stream Control Transmission Protocol}
\newacro{SDP}{Session Description Protocol}
\newacro{SigOfly}{Signaling-On-the-fly}
\newacro{SIMPLE}{SIP for Instant Messaging and Presence Leveraging Extensions}
\newacro{SIP}{Session Initiation Protocol}
\newacro{SMIL}{Synchronized Multimedia Integration Language}
\newacro{SMS}{Short Message Service}
\newacro{SRTCP}{Secure RTCP}
\newacro{SRTP}{Secure Real-time Transport Protocol}
\newacro{SRT}{SubRip Text}
\newacro{SSL}{Secure Sockets Layer}
\newacro{STUN}{Session Traversal Utilities for NAT}
\newacro{SVG}{Scalable Vector Graphics}
\newacro{TCP}{Transmission Control Protocol}
\newacro{TLS}{Transport Layer Security}
\newacro{TMN}{This Means Nothing}
\newacro{TURN}{Traversal Using Relays around NAT}
\newacro{UDP}{User Datagram Protocol}
\newacro{VoIP}{Voice Over IP}
\newacro{WebRTC}{Web Real-Time Communication}
\newacro{WebVTT}{Web Video Text Tracks}
\newacro{XAML}{eXtensible Application Markup Language}
\newacro{XEP}{XMPP Extensions}
\newacro{XHTML}{eXtensible Hypertext Markup Language}
\newacro{XML}{eXtensible Markup Language}
\newacro{XMPP}{Extensible Messaging and Presence Protocol}

%\overfullrule=2cm

\newganttchartelement*{mymilestone}{
mymilestone/.style={
shape=isosceles triangle,
inner sep=0pt,
draw=cyan,
top color=white,
bottom color=cyan!50
},
mymilestone incomplete/.style={
/pgfgantt/mymilestone,
draw=yellow,
bottom color=yellow!50
},
mymilestone label font=\slshape,
mymilestone left shift=0pt,
mymilestone right shift=0pt
}

\newgantttimeslotformat{stardate}{%
\def\decomposestardate##1.##2\relax{%
\def\stardateyear{##1}\def\stardateday{##2}%
}%
\decomposestardate#1\relax%
\pgfcalendardatetojulian{\stardateyear-01-01}{#2}%
\advance#2 by-1\relax%
\advance#2 by\stardateday\relax%
}

\begin{document}
\pagestyle{plain}
\mainmatter              % start of the contributions

\title{Hyper-linked Communications: WebRTC enabled asynchronous collaboration}
\author{%
	Henrique Duarte Lopes Rocha \\
	email: henrique.rocha@tecnico.ulisboa.pt \\
  student number: 68621
}
\institute{Master Degree in Telecommunications and Informatics Engineering \\ Instituto Superior Técnico \\ Universidade de Lisboa}
%RP acrescenta número de aluno, nome do mestrado e ``cadeira de projecto''
\maketitle              % typeset the title of the contribution

\begin{abstract}
The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content, to communication and collaboration services. This paradigm allows to synchronize, structure and navigate communication content, providing a way to integrate social media content and collaborative tools into voice and video calls.
Voice and image together can express emotions and expose creativity like no other medium can. With hypermedia concepts, we can add more value to voice and video communications.

WebRTC technology allows real-time communications between web browsers without the need to install and use additional applications or plug-ins. The nature of web browser applications already follow the hypermedia concept, which makes WebRTC the ideal technology to apply the hyper-linked communications concepts.
The native support of WebRTC in operating systems extends its usage to outside the web browser, allowing to explore functionalities that browsers are poor to support such as video recording and massive information storage.

Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable colaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed latter.
This application will target the web browser platform, resorting to WebRTC.

In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose an architecture to implemente an hyper-linked communication application based on WebRTC and provide a plan for its implementation and evaluation.

\keywords{WebRTC, asynchronous, communications, collaboration}
\end{abstract}

\section{Introduction}
  \input{intro.tex}

\section{Related Work}\label{related}
 This section is structured as follows.
 Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. 
 Section \ref{rtc} describes the \ac{WebRTC} technology and the protocols needed to implement our project. 
 Section \ref{signaling} addresses the signaling component of chat applications, which is not defined on \ac{WebRTC} specifications. 
 Section \ref{hypermedia} presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity. 
 Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
%RP este detalhe sobre a estrutura da sec. do estado da arte vai para o início dessa secção.

  \input{state01.tex}
  \input{state02.tex}
  \input{state03.tex}
  \input{state04.tex}

\section{Proposed Architecture}\label{arch} % English
  \input{arch01.tex}

\section{Methodology}\label{meth} % Section

		\input{method.tex}

	\subsection{Planned Schedule} % English
		\input{schd01.tex}

\section{Conclusions}\label{concl} % English

%\subsection{Summary} % English
%Neste secção deve-se fazer o resumo do trabalho efectuado, retomando a ideia, as contribuições definidas e a forma como estas se materializaram

%\subsection{Conclusions} % English
%Nesta secção devem ser retiradas conclusões do trabalho realizado, em face dos resultados obtidos \cite{exemplo}.

\bibliographystyle{splncs03}
\bibliography{references}

\end{document}