### 2016-01-21 curriculo/grades_master.tex

\textit{Mobile, Peer-To-Peer and Cloud Applications}
	\begin{itemize}
		\item Cloud Computing - 15
		\item Mobile Computing - 15
		\item Peer-to-Peer Systems and Overlay Networks - 15
	\end{itemize}
\textit{Entrepreneurial Leadership, Organization, Economy and Management}
	\begin{itemize}
		\item Decision Support Models - 14
	\end{itemize}
\textit{Common Branch}
	\begin{itemize}
		\item Forensics Cyber-Security - 16
		\item Middleware for Distributed Internet Applications - 14
		\item Telecommunications and Positioning Satellite Systems - 17
		\item Performance Evaluation and Dimensioning of Networks and Systems - 15
		\item Web Data Extraction and Analysis - 16
	\end{itemize}
\textit{Master Thesis}
	\begin{itemize}
		\item Master Project in Telecommunications and Computer Engineering - 17
	\end{itemize}

\textit{Elective Courses}
	\begin{itemize}
		\item Services Engineering - 17
		\item Database Administration and Tuning - 15
	\end{itemize}