### 2016-02-21 project/stats/draft.tex
[ADD] stats folder

\documentclass[
10pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
]{scrartcl}

\usepackage{amsmath} 
\usepackage{adjustbox}
\usepackage{textcomp}
\usepackage{environ}
\usepackage{tikz}
\usepackage{graphics}
\usepackage{pgfplots} 
\pgfplotsset{compat=1.9}

\title{Draft for master thesis} % The article title

\author{Henrique Rocha} 
\date{} 
\begin{document}

	\maketitle % Print the title/author/date block

	%----------------------------------------------------------------------------------------
	%	INTRODUCTION
	%----------------------------------------------------------------------------------------

	\section{Test}

\subsection{Usability Test}

	\subsubsection{Task 1 (Video)}
		Register User, add the coordinator as a friend, create a room, add the coordinator to the room wait for the coordinator to join the room and change for the individual view.

	\subsubsection{Task 2 (Tags)}
		The coordinator creates previously a room with existing content, namely, tags and subtitles.
		The user task is to enter in the public room and search by "Cauchy Rule" get an url for that time and send it back to the other users.
		Next the user should create a tag on exact arrival time (when he entered the room).

	\subsubsection{Task 3 (Content)}
		The user must create a subtitle with 10s duration at the time of arrival with the content "I've just arrived" and search for the previously created tag

	\subsubsection{Task 4 (Colaborative)}

\subsection{Performance Test}

	\begin{center}
	\input{points3.tex}
	\end{center}

	\begin{center}
	\input{points2.tex}
	\end{center}

	\begin{center}
	\input{points.tex}
	\end{center}
\end{document}