### 2016-03-12 tese1/04_goals.tex

\subsection{Thesis Goals} % English

	Making it clear, this project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (e.g. creating annotations for improving content search) in order to increase its value. Another important goal of this project is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.

A web application with an easy to learn user interface will be developed to accomplish solving our problem. Our application, unnamed yet, is targeted at web browsers that are compatible with only standard technologies like JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3. Any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.

%RP podes elaborar um pouco. Objectivos são: realizar o levantamento do estado da arte no espaço do problema para fundamentar uma solução; apresentar uma solução e desenhar uma arquitectura para dar resposta ao problema apresentado; construir protótipo funcional; desenhar e realizar uma avaliação/testes que provem a adequação e validade dos conceitos desenvolvidos (elaborar); escrever um artigo.

Although we propose an initial solution to solve our problem, we will make a continuous effort to survey the systems that solve our problem either completely or just a part of it, not excluding solutions that may appear during the implementation of our project.

We will present an architecture that can meet our goals, implement the respective prototype and test it with real users, unitary tests and benchmarks.

According to Martin Geddes, the quality of the interaction worsens as the number of users increase\cite{geddes}. In our testing phases we will quantify and qualify the impact of increasing users on the interface and performance of our prototype.

	All the problems faced during the development and limitations will be reported on the thesis so that a future project better then ours can be easily and better developed.
        %RP esta última frase não é muito necessária