### 2016-03-12 tese1/project.tex

\documentclass{llncs}
\RequirePackage[english]{babel} % Se fizerem o texto em ingles
\usepackage[english]{babel}   
%\usepackage[latin1]{inputenc}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

\usepackage[fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{subfig}  % allows for indexgeneration
\usepackage{color}  % allows for indexgeneration
\usepackage{url}
\usepackage{hyperref}
\usepackage{float}
\usepackage{mathtools}
\usepackage{acronym}
\usepackage{pgfgantt}
\usepackage{adjustbox}
\usepackage{rotating}
\usepackage{dirtytalk}
\usepackage[graphicx]{realboxes}
\definecolor{fade}{rgb}{0.8,0.8,0.8}

\setcounter{secnumdepth}{3}
\newacro{AES}{Advanced Encryption Standard}
\newacro{API}{Application Programming Interface}
\newacro{BOSH}{Bidirectional-streams Over Synchronous HTTP}
\newacro{CBC}{Cipher Block Chaining}
\newacro{CPU}{Central Processing Unit}
\newacro{CRUD}{Create, Read, Update and Delete}
\newacro{CSS}{Cascading Style Sheets}
\newacro{DOM}{Document Object Model}
\newacro{DTLS}{Datagram Transport Layer Security}
\newacro{DVD}{Digital Versatile Discs}
\newacro{FMS}{Flash Media Server}
\newacro{HTML}{HyperText Markup Language}
\newacro{HTTP}{Hypertext Transfer Protocol}
\newacro{ICE}{Interactive Connectivity Establishment}
\newacro{IPv4}{Internet Protocol Version 4}
\newacro{IPv6}{Internet Protocol Version 6}
\newacro{IP}{Internet Protocol}
\newacro{JSON}{JavaScript Object Notation}
\newacro{MPEG}{Moving Picture Experts Group}
\newacro{MVC}{Model-View-Controller}
\newacro{NAT}{Network Address Translation}
\newacro{OT}{Operational transformation}
\newacro{P2P}{Peer-to-peer}
\newacro{PSTN}{Public Switched Telephone Network}
\newacro{RFC}{Request For Comments}
\newacro{RTCP}{Real Time Control Protocol}
\newacro{RTSP}{Real Time Streaming Protocol}
\newacro{RTMFP}{Real-Time Media Flow Protocol}
\newacro{RTMP}{Real Time Messaging Protocol}
\newacro{RTP}{Real-time Transport Protocol}
\newacro{SAMI}{Synchronized Accessible Media Interchange}
\newacro{SCTP}{Stream Control Transmission Protocol}
\newacro{SDP}{Session Description Protocol}
\newacro{SigOfly}{Signaling-On-the-fly}
\newacro{SIMPLE}{SIP for Instant Messaging and Presence Leveraging Extensions}
\newacro{SIP}{Session Initiation Protocol}
\newacro{SMIL}{Synchronized Multimedia Integration Language}
\newacro{SMS}{Short Message Service}
\newacro{SRTCP}{Secure RTCP}
\newacro{SRTP}{Secure Real-time Transport Protocol}
\newacro{SRT}{SubRip Text}
\newacro{SSL}{Secure Sockets Layer}
\newacro{STUN}{Session Traversal Utilities for NAT}
\newacro{SVG}{Scalable Vector Graphics}
\newacro{TCP}{Transmission Control Protocol}
\newacro{TLS}{Transport Layer Security}
\newacro{TMN}{This Means Nothing}
\newacro{TURN}{Traversal Using Relays around NAT}
\newacro{UDP}{User Datagram Protocol}
\newacro{VoIP}{Voice Over IP}
\newacro{WebRTC}{Web Real-Time Communication}
\newacro{WebVTT}{Web Video Text Tracks}
\newacro{XAML}{eXtensible Application Markup Language}
\newacro{XEP}{XMPP Extensions}
\newacro{XHTML}{eXtensible Hypertext Markup Language}
\newacro{XML}{eXtensible Markup Language}
\newacro{XMPP}{Extensible Messaging and Presence Protocol}

%\overfullrule=2cm

\newganttchartelement*{mymilestone}{
mymilestone/.style={
shape=isosceles triangle,
inner sep=0pt,
draw=cyan,
top color=white,
bottom color=cyan!50
},
mymilestone incomplete/.style={
/pgfgantt/mymilestone,
draw=yellow,
bottom color=yellow!50
},
mymilestone label font=\slshape,
mymilestone left shift=0pt,
mymilestone right shift=0pt
}

\newgantttimeslotformat{stardate}{%
\def\decomposestardate##1.##2\relax{%
\def\stardateyear{##1}\def\stardateday{##2}%
}%
\decomposestardate#1\relax%
\pgfcalendardatetojulian{\stardateyear-01-01}{#2}%
\advance#2 by-1\relax%
\advance#2 by\stardateday\relax%
}

\begin{document}
\pagestyle{plain}
\mainmatter              % start of the contributions

\title{Hyper-linked Communications: WebRTC enabled asynchronous collaboration}
\author{%
	Henrique Duarte Lopes Rocha \\
	email: henrique.rocha@tecnico.ulisboa.pt \\
  student number: 68621
}
\institute{Master Degree in Telecommunications and Informatics Engineering \\ Instituto Superior Técnico \\ Universidade de Lisboa}
%RP acrescenta número de aluno, nome do mestrado e ``cadeira de projecto''
\maketitle              % typeset the title of the contribution

\input{01_abstract.tex}
\input{02_introduction.tex}
\input{03_problem.tex}
\input{04_goals.tex}
\input{05_structure.tex}
\input{06_related_work.tex}
\input{07_related_work_nat.tex}
\input{08_related_work_rtc.tex}
\input{09_related_work_signaling.tex}
\input{10_related_work_hypermedia.tex}
\input{11_related_work_collaboration.tex}
%\input{XX_implementation_model.tex}

\section{Proposed Architecture}\label{arch} % English
  \input{arch01.tex}

\section{Methodology}\label{meth} % Section

\input{method.tex}

	\subsection{Planned Schedule} % English
		\input{schd01.tex}

\section{Conclusions}\label{concl} % English
The development of \ac{WebRTC} by the World Wide Web Consortium (W3C) allows web browser-to-browser audio, video and data communication. \ac{WebRTC} is enabling new usage scenarios for communication and collaboration applications. This and other technologies allow the Hyper-link concept to be applied to a new media: real-time audio and video.

  This project proposes a solution that uses \ac{WebRTC} for creating an application that provides audio, data, text and data real-time communications. Theses communications will be enriched using hypermedia concepts, the possibility to record and playback its content and the integration of collaborative tools.

  A prototype will be implemented in order to validate these concepts.
In order to implement it, significant challenges, such as the need to record interactive audio and video streams, will have to be overcome, making this a challenging project.

\bibliographystyle{splncs03}
\bibliography{references}

\end{document}