### 2016-03-14 tese1/XX_implementation_signaling.tex
[RFT] improved protocol readability

\subsection{Signaling Protocol}

Our signaling protocol consists in sending JSON formated messages over WebSockets