### 2016-03-19 curriculo/cv.tex
[]

%!TEX TS-program = xelatex
\documentclass[]{friggeri-cv}
\addbibresource{bibliography.bib}
\usepackage{listings}
\usepackage{color}

\definecolor{pblue}{rgb}{0.13,0.13,1}
\definecolor{pgreen}{rgb}{0,0.5,0}
\definecolor{pred}{rgb}{0.9,0,0}
\definecolor{pgrey}{rgb}{0.46,0.45,0.48}

\lstset{language=Java,
  showspaces=false,
  showtabs=false,
  breaklines=true,
  showstringspaces=false,
  breakatwhitespace=true,
  commentstyle=\color{pgreen},
  keywordstyle=\color{pblue},
  stringstyle=\color{pred},
  basicstyle=\ttfamily,
  moredelim=[il][\textcolor{pgrey}]{$$},
  moredelim=[is][\textcolor{pgrey}]{\%\%}{\%\%}
}

\begin{document}
\header{Henrique}{Rocha}
       {Telecommunications and Informatics}

% In the aside, each new line forces a line break
\begin{aside}
  \section{about}
    Henrique Duarte Lopes Rocha
    ~
    Rua dos Cedros nº100
    Alto do Lagoal
    2760-025 Caxias
    Portugal
    ~
    (+351) 938781922
    ~
    \href{mailto:hdlopesrocha91@gmail.com}{hdlopesrocha91@gmail.com}
    \href{http://web.ist.utl.pt/ist168621}{WebSite}
    \href{http://www.youtube.com/user/hdlopesrocha}{YouTube}
  \section{languages}
    portuguese (native)
    english (fluent)
  \section{programming}
    {\color{red} $\varheartsuit$} Java
	 (Android, PlayFramework, LWJGL, MapReduce, GWT, AWS, JavaFX, Kurento Media Server, Jetty and Jersey), C, C++ (OpenGL), C\# (XNA, Windows Forms, WPF, lpsolve), VB, Python, Scala, Scheme, HTML, CSS (Bootstrap), JavaScript (jQuery, AJAX, WebGL, WebRTC), PHP (Wordpress), GLSL, HLSL.  
   \section{databases}
	DynamoDB, MongoDB, MySQL, Neo4j, PostgreSQL, SQLite.
   \section{soft skills}
	Time Management
	Problem-Solving
	Team Player
	Self Confidence
	Adaptability
\end{aside}

\section{education}

\begin{entrylist}
   \entry
    {2013–now}
    {Master of Science}
    {Instituto Superior Técnico - Taguspark}
    {Telecommunications and Informatics Engineering (Expected Grade: 16 in 20)}
   \entry
    {2009–2014}
    {Bachelor of Science}
    {Instituto Superior Técnico - Taguspark}
    {Communication Networks Engineering (Final Grade: 13.2 in 20)}

   \entry
    {2006–2009}
    {Secondary School}
    {Escola Secundária Luís de Freitas Branco, Paço de Arcos}
    {Science \& Technology}

\end{entrylist}

\section{experience}

\begin{entrylist}

  \entry
    {01 2016\\(ongoing)}
    {Applications Development}
    {Bullray-CIT}
    {\emph{Development of a user interface for a bank's security device that is synchronized with servers. Microcontroller sensors information receival through \underline{USB} (\underline{org.usb4java}), user interface with \underline{JavaFX} and network camera discovery through nmap (using \underline{org.nmap4j}) } }

 \entry
    {11 2015\\(2 months)}
    {Android Development}
    {Bullray-CIT}
    {\emph{Optimization and bug resolution of an \underline{Android} application that uploads GPS, camera and microphone data in background.}}

\entry
    {08 2015\\(standby)}
    {Systems Development}
    {Bullray-CIT}
    {\emph{Development and configuration of a streaming server that supports receiving and recording video from authenticated users through \underline{ffmpeg}. I tried \underline{FFServer}, \underline{nginx with rtmp module} and \underline{Wowza Streaming Server}, the final decision was \underline{Kurento Media Server} with an \underline{RTP} endpoint for receiving video from network cameras and a \underline{WebRTC} endpoint for delivering the video to web browsers. }}

\entry
    {03 2015\\(ongoing)}
    {Master Thesis}
    {IST}
    {\emph{Development of an interactive multi user video chat that supports recording, content overlay and a collaborative text editor. Used \underline{Kurento Media Server} for mixing streams into a single one and detecting QR codes. Experience with \underline{ot.js} for operation transformations (collaborative component). Other libraries used, \underline{jquery.qrcode}, \underline{vis.js}, \underline{bootstrap-typeahead}, \underline{codemirror.js} and \underline{adapter.js} }. Little experience with \underline{strophe.js} for communication between web browsers and \underline{XMPP} servers. }

  \entry
    {02 2015\\(ongoing)}
    {Full-Stack Web Development}
    {Bullray-CIT}
    {\emph{Development of a solution for a generic incident handling server, support for real time monitoring through \underline{WebSockets}. Experience with \underline{PlayFramework}, \underline{MongoDB} (with the official Java Driver), \underline{Maven}, \underline{jQuery} and \underline{Bootstrap}.}}

  \entry
    {11 2014\\(4 months)}
    {Full-Stack Web Development}
    {Bullray-CIT}
    {\emph{Development of an architecture for a chat application. Experience with \underline{Android}, \underline{PlayFramework}, \underline{MongoDB} (with Morphia), \underline{jQuery} and \underline{Bootstrap}. Message synchronization through HTTP long polling.}}
  \entry
    {04 2014\\(6 months)}
    {Scientific Initiation Scholarship}
    {INESC}
    {\emph{Frontend Development for Provide Results's Bimk project. Experience with \underline{PlayFramework}, \underline{Revit API} and \underline{GIT}.}}
  \entry
    {09 2013\\(6 months)}
    {Scientific Initiation Scholarship}
    {CEG-IST}
    {\emph{Programming a Decision Analysis Software (MACBETH) with \underline{WPF}. Development of an algorithm to suggest judgements based on past decisions using \underline{lpsolve}.}}
  \entry
    {08 2013\\(2 weeks)}
    {Summer Work}
    {LCG}
    {\emph{Frontend development for a Decision Analysis Software using \underline{ASP.NET}. }}
  \entry
    {11 2012}
    {Pizza Night Competition}
    {Microsoft Lisbon Experience}
    {\emph{2º place on Windows Phone Category} with \emph{Infinity Edge} game.}
  \entry
    {08 2006\\(1 month)}
    {Garbage Collector}
    {Jovens em Movimento, Oeiras}
    {Cleaning and maintenance of public spaces.}

\end{entrylist}
%\newpage

\section{applications}

\begin{entrylist}
%  \entry
%    {201?}
%    {My personal website}
%    {\href{http://play.google.com/store/apps/details?id=pt.ist.bombraider}{henrique.rocha.pt}}
%    {I like PlayFramework, but I still like more to learn different technologies, that's why I did this website using ???, just for fun.}

  \entry
    {2015}
    {Bomb Raider}
    {\href{http://play.google.com/store/apps/details?id=pt.ist.bombraider}{Google Play}}
    {Bomb Raider is a simple singleplayer and multiplayer game where the goal is destroying all enemies in the map. I used \underline{OpenGLES} and \underline{WiFi-Direct} to develop this \underline{Android} game.}

  \entry
    {2013}
    {OpenGlobe}
    {\href{http://web.ist.utl.pt/ist168621/?page=10}{Instituto Superior Técnico}}
    {Modeling Earth surface with GPS devices and topographic collected data. It is also possible to predict satellites position through ephemeris data.}

  \entry
    {2012}
    {Infinity Edge (Windows Phone 7)}
    {\href{http://www.windowsphone.com/en-us/store/app/infinity-edge/05516463-3a89-4351-996e-62e5b4519aeb}{windows marketplace}}
    {First Person Shooter with spaceships made on my free time. I had experience with \underline{XNA} and \underline{HLSL}.}

  \entry
    {2011}
    {Era of Empires}
    {\href{http://web.ist.utl.pt/ist168621/?page=8}{Personal Project}}
    {3D environment generator and Real Time Strategy game engine using \underline{OpenGL}.}

\end{entrylist}

\section{interests}
    3D Studio Max, Photoshop, Game Development, Web Design, Network Security, Jogging, Karate and Yoga.

%	My mother said I should have a girlfriend...

%\newpage
%\section{grades}
%\subsection{Bachelor - Communication Networks Engineering}
%\input{grades_bachelor.tex}
%\subsection{Master - Telecommunications and Informatics Engineering}
%\input{grades_master.tex}

\begin{lstlisting}
return 0;
\end{lstlisting}
\end{document}