### 2016-03-25 tese2/architecture.tex

\chapter{Architecture}
\label{chapter:architecture}
%Esta secção deve englobar:

%\begin{itemize}
%    \item Uma análise de requisitos, onde devem “abrir” portas para as principais
%        características da vossa arquitectura.
%    \item Uma descrição de alto nível da arquitectura do vosso sistema. Esta descrição
%        pode ser:
%        \begin{itemize}
%            \item A arquitectura de rede, explicando os principais componentes e as
%                funções que executam.
%            \item A arquitectura de SW da aplicação, explicando os principais
%                componentes e as funções que executam.
%        \end{itemize}
%    \item Descrições detalhadas de todos os componentes da vossa arquitectura.
%    \item As escolhas que efectuarem em termos de arquitectura global e de detalhe,
%        devem ser justificadas, face aos requisitos que identificaram.
%\end{itemize}

%Sempre que possível, ilustrem a arquitectura com figuras, que descrevam
%sucintamente o modelo como na Figure~\ref{fig:logoCNM}.

%\begin{figure}[!htb]
%  \centering
%  \includegraphics[width=0.25\textwidth]{Figures/logoCNM.png}
%  \caption[Caption for figure in TOC]{Caption for figure.}
%  \label{fig:logoCNM}
%\end{figure}

Taking into account the goals of this project and all the technology presented so far. Our proposal is the development of a web application that provides communication and collaboration in real-time.

The requirements for our web application are:

\begin{itemize}
 \item Text, Audio and Video communication using \ac{WebRTC}.
 \item Ability to record and playback interactive multimedia.
 \item Create annotations over video and voice content.
 \item Enrich communications by overlaying multiple media types.
 \item Search and navigate through annotations and contents.
 \item Ability to create and share time links among users.
 \item Mix multiple user streams into a single stream.
 \item Sound detection for showing the current speaker.
 \item Structure presentation of the content.
 \item Collaborative text edition.
 \item Easy interface for content creation and synchronization. 
\end{itemize}

	Clients can discover chat rooms and other clients by querying the signaling server, they can create rooms for multi-party communication which is achieved by using \ac{WebRTC}'s \emph{MediaStream} and \emph{DataChannel}, namely for audio, video and text communication.

	Chat rooms can be public or private. Public chat rooms are moderated by a group of clients which initially is formed by the room creator, this type of room has no access restrictions by default, but that can be changed by its moderators. Private chat rooms will be visible for a list of clients or can be accessed by clients that have a link for that chat room.

\section{Modules}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\textwidth]{figures/modules.pdf}
	\caption{System Modules}
        \label{fig:modules}
\end{figure}
%RP as figuras têm de ser apresentadas. Figure~\ref{fig:modules} presents ...
	Figure~\ref{fig:modules} presents the structure of our system which is divided into six modules:

\begin{itemize}

\item \underline{Application module} - responsible to provide information about the relevant modules (\emph{NAT Traversal} and \emph{Signaling}) and user interface to the \emph{Client}.

\item \underline{Signaling module} - responsible for \emph{Client} and \emph{Stream} coordination.

 \item \underline{NAT Traversal module} - used by \emph{Client} and \emph{Stream} modules during the \emph{Signaling} phase which ends by establishing the connection between them.

 \item \underline{Stream module} - responsible to deliver and receive multimedia content from the \emph{Client}.

 \item \underline{Storage module} - responsible store all the model information and media recorded.

 \item \underline{Client module} - responsible for the interaction with the user.

\end{itemize}

% Each room will be associated to multiple Media-Types, interactive and non-interactive, discrete and continuous. This media types can be structured with chains of events. Some components of the user interface may not be synchronized to all participants, for example an help or suggestion window.

\section{Implementation Proposal}
The infrastructure is composed by: Ice Server, Web Server, Stream Server, Signaling Server and Test Clients.

\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{figures/infrastructure.pdf}
	\caption{System Infrastructure}
\end{figure}

In order to simplify our solution, we propose both the \emph{Application} and \emph{Signaling} modules implementation within the same server application so it could be easy to control which resources the client has permission to access.

We have taken into account the compatibility between the streaming server, database and the operation transformation solutions in order choose the appropriate framework to implement our web server.

Based on our study about signaling protocols we decided to exclude \ac{XMPP} due to the development difficulties that it exposes when using multiple web pages, namely the user re-authentication performed each time a page is loaded and the single tab usage limitation. By consequence we have not choosen \emph{Jitsi Video Brigde} as it uses \ac{XMPP}.

By excluding \emph{Jitsi Video Brigde} we have decided that our solution must use \ac{KMS}. Our web server could be implemented more easily with \emph{NodeJS} or \emph{Java} due to the fact \ac{KMS} provides clients for both technologies but others could be used as \ac{KMS} also exposes their \ac{API} via \emph{WebSockets}.

By default, \emph{Kurento Repository} is implemented over \emph{MongoDB}, for convenience our storage model will also use the same database.

For the web server we could choose \emph{NodeJS} or any \emph{Java} based web framework. One of our criteria for choosing the web framework is the ability to follow the \ac{MVC} paradigm which can help us to organize our code. Without any strong reason we have decided to implement our webserver with \emph{PlayFramework}\footnote{\url{https://www.playframework.com/}(acessed March 25, 2016)} using \emph{Java}.

For the collaborative text editor, we have choosen \emph{OT.js} due to its indenpendence from server implementation and storage choice.

For the \emph{NAT Traversal} module an \ac{ICE} server is not required to be on our infrastructure, as a public \ac{STUN} server can be used for testing our solution.

On the client computers, both \emph{Mozilla Firefox} and \emph{Google Chrome} should be installed as web browsers. Libraries such as \emph{jQuery}, \emph{Bootstrap}, \emph{Adapter.js}, \emph{OT.js} can be downloaded from the Web Server and executed on the client side.

\begin{table}[H]
\centering

    \begin{tabular}{cccccccc@{}m{0pt}@{}}
	\hline 
\multicolumn{8}{|c|}{\cellcolor{Gray}Application}  &\\[12pt]\cline{1-5}\cline{7-7}
\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{ot.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} & \multicolumn{1}{c|}{adapter.js} &   \multicolumn{1}{c|}{\cellcolor{Gray}} &\\[12pt]\hline
\multicolumn{1}{|c|}{HTTP} & \multicolumn{2}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[12pt]\hline
\end{tabular}
	\caption{App Architecture}
	        \label{fig:apparch}
\end{table}

Figure~\ref{fig:apparch} presents the application architecture and the underlying technologies. \emph{Adapter.js} and \emph{jQuery} will ensure that our application is compatible with the most popular web browsers.
\emph{Bootstrap} will be used to make the user interface more appellative and responsive. With \emph{Bootstrap} it is quite easy to develop applications that adapt to mobile devices with different screen sizes.

The synchronization between multimedia elements will be performed trough chains of \emph{JavasSript} events or by specifying the interval of time which time content must be visible. Other animations can be implemented with \ac{SVG} embedded on \ac{HTML}.