### 2016-03-26 project/stats/true_cpu.tex

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Cpu Usage},
	    xlabel={Time [seconds]},
	    ylabel={CPU [percentage]},
	    xmin=0, xmax=800,
	    ymin=0, ymax=1200,
	    xtick={0,100,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={10,70,...,1000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_7_mongod.csv};
		\addlegendentry{Mongod}

		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_67_kurento-media-server.csv};
		\addlegendentry{Kurento Media Server}

		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_35_kurento-repo.csv};
		\addlegendentry{Kurento Repository}

		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_233_bash.csv};
		\addlegendentry{Java}

	\end{axis}
\end{tikzpicture}