### 2016-03-27 tese2/stats/summary_full_net.tex
[ADD] amazing explanation of network test results

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=8cm,
	    title={Network usage},
	    xlabel={Time [seconds]},
	    ylabel={Transfer rate [kB/s]},
	    xmin=-10, xmax=790,
	    ymin=0, ymax=2000,
	    xtick={0,60,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

	\addplot[color=blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_sent_eth0.csv};
	\addlegendentry{Sent(eth0)}

	\addplot[color=red] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_recv_eth0.csv};
	\addlegendentry{Recv(eth0)}

	\addplot[color=green] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_sent_lo.csv};
	\addlegendentry{Sent/Recv(lo)}

	\end{axis}
\end{tikzpicture}