### 2016-04-07 tese2/stats/test_client_net.tex
[ADD] client side tests

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Network usage},
	    xlabel={Time [seconds]},
	    ylabel={Transfer rate [kbps]},
	    xmin=-10, xmax=790,
	    ymin=0, ymax=1000,
	    xtick={0,60,...,2000},
	    ytick={0,100,...,5000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

	\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-0,y expr=\thisrowno{1}/1000] {stats/client/send.dat};
	\addlegendentry{Sent(eth0)}

	\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-0,y expr=\thisrowno{1}/1000] {stats/client/recv.dat};
	\addlegendentry{Recv(eth0)}

	\end{axis}
\end{tikzpicture}