### 2016-04-17 tese2/stats/test_two_times_mem.tex

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=8cm,
	    title={Memory Usage},
	    xlabel={Time [seconds]},
	    ylabel={Memory [MB]},
	    xmin=-10, xmax=1690,
	    ymin=0, ymax=1800,
	    xtick={0,120,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

		\addplot[color=Blue] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_7_mongod.dat};
		\addlegendentry{Mongod}

		\addplot[color=Red] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_68_kurento-media-server.dat};
		\addlegendentry{Kurento Media Server}

		\addplot[color=Green] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_35_kurento-repo.dat};
		\addlegendentry{Kurento Repository}

		\addplot[color=Magenta] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_234_bash.dat};
		\addlegendentry{Java}

	\end{axis}
\end{tikzpicture}