### 2016-04-23 tese2/stats/old/test_hour_cpu.tex

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=8cm,
	    title={CPU usage},
	    xlabel={Time [seconds]},
	    ylabel={CPU [percentage]},
	    xmin=0, xmax=3600,
	    ymin=0, ymax=1000,
	    xtick={0,500,...,5000},
	    ytick={0,100,...,2000},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

		\addplot[color=Blue,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_8_mongod.dat};
		\addlegendentry{Mongod}

		\addplot[color=Red,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_91_kurento-media-server.dat};
		\addlegendentry{Kurento Media Server}

		\addplot[color=Green,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_36_kurento-repo.dat};
		\addlegendentry{Kurento Repository}

		\addplot[color=Magenta,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_228_bash.dat};
		\addlegendentry{Java}

	\end{axis}
\end{tikzpicture}