### 2016-04-23 tese2/stats/old/test_hour_net.tex

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=8cm,
	    title={Network usage},
	    xlabel={Time [seconds]},
	    ylabel={Transfer rate [mbps]},
	    xmin=0, xmax=3600,
	    ymin=0, ymax=24,
	    xtick={0,500,...,5000},
	    ytick={0,5,...,100},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

	\addplot[color=Blue,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}*8/1000000] {stats/test_hour/sent_eth0.dat};
	\addlegendentry{Sent(eth0)}

	\addplot[color=Red,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}*8/1000000] {stats/test_hour/recv_eth0.dat};
	\addlegendentry{Recv(eth0)}

	\addplot[color=Green,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}*8/1000000] {stats/test_hour/recv_lo.dat};
	\addlegendentry{Sent/Recv(lo)}

	\end{axis}
\end{tikzpicture}