### 2016-04-30 tese2/acknowledgements.tex

\chapter*{\acknowledgments}

% Add entry in the table of contents as chapter
\addcontentsline{toc}{chapter}{\acknowledgments}

I would like to thank my dissertation supervisor Ricardo Pereira and co-supervisor Paulo Chainho for their active support and ideas that were crucial for the development of this thesis.

I should not forget all the people involved on the prototype's test phase, they not only tested but also gave valuable feedback that we used to improve the quality of our prototype.

Also, I would like to show my gratitude to the \emph{Kurento} team for developing the streaming server that I've used, their impressive help support and receptiveness to discuss new functionalities.

Not less important I should thank all IST employees that maintained the space I used during this journey.

Lastly, I should thank all the people that supported me psychologically and made me going my own way.

Thank you!

\cleardoublepage