### 2016-05-06 tese2/architecture.tex

\chapter{Architecture}
\label{chapter:architecture}
%Esta secção deve englobar:

%\begin{itemize}
%    \item Uma análise de requisitos, onde devem “abrir” portas para as principais
%        características da vossa arquitectura.
%    \item Uma descrição de alto nível da arquitectura do vosso sistema. Esta descrição
%        pode ser:
%        \begin{itemize}
%            \item A arquitectura de rede, explicando os principais componentes e as
%                funções que executam.
%            \item A arquitectura de SW da aplicação, explicando os principais
%                componentes e as funções que executam.
%        \end{itemize}
%    \item Descrições detalhadas de todos os componentes da vossa arquitectura.
%    \item As escolhas que efectuarem em termos de arquitectura global e de detalhe,
%        devem ser justificadas, face aos requisitos que identificaram.
%\end{itemize}

%Sempre que possível, ilustrem a arquitectura com figuras, que descrevam
%sucintamente o modelo como na Figure~\ref{fig:logoCNM}.

%\begin{figure}[!htb]
%  \centering
%  \includegraphics[width=0.25\textwidth]{Figures/logoCNM.png}
%  \caption[Caption for figure in TOC]{Caption for figure.}
%  \label{fig:logoCNM}
%\end{figure}

Taking into account the goals of this project and all the technology presented so far. Our proposal is the development of a web application that provides communication and collaboration features in real-time.

%RP isto pode ser muito expandido. Podes começar por descrever os goals (mais alto nível): audio e video conferência multi-party, possibilidade de rever o que foi dito antes, chat, etc etc.
% Só depois traduzes isso em requisitos, tanto funcionais como não funcionais
% Depois apresentas a arquitectura, fazendo um paralelo entre os módulo e escolhas realizadas e quais os requisitos que cada um deles satisfaz.

The requirements for our web application are:

\begin{itemize}
 \item Text, Audio and Video communication using \ac{WebRTC}.
 \item Ability to record and playback interactive multimedia.
 \item Create annotations over video and voice content.
 \item Enrich communications by overlaying multiple media types.
 \item Search and navigate through annotations and contents.
 \item Ability to create and share time links among users.
 \item Mix multiple user streams into a single stream.
 \item Sound detection for showing the current speaker.
 \item Structure presentation of the content.
 \item Collaborative text edition.
 \item Easy interface for content creation and synchronization. 
\end{itemize}

%RP capacidade de lidar com NAT

	Moreover we allow clients to discover chat rooms and other clients by navigating on the web pages provided by our web server. In addition users can create rooms for multi-party audio and video communication communication which is achieved by using \ac{WebRTC}'s \emph{MediaStream} and \emph{PeerConnection}.
        %RP falas em peerconnection, mas na figura 3.1 apenas há um cliente, o que obviamente impede mostrar a comunicação entre peers.
        %RP acho que podes fazer outro diagrama com a comunicação entre os vários componentes, onde apareceriam vários clientes. A aí indicas o tipo de ligação entre cada par de componentes (HTTP, PeerConnectino, websocket, MediaStream ...)

	Furthermore, chat rooms can be public or private. Public chat rooms are moderated by a group of clients which initially consistes only of the room creator. This type of room has no access restrictions by default, but that can be changed by its moderators. Private chat rooms will only be visible to a defined list of clients or can be accessed by clients that have a link for that conference room.

\section{Modules}
In this section we present the several modules that were designed in order com fulfill the set requirements.
	Figure~\ref{fig:modules} presents the structure of our system which was divided into six modules:

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.8\textwidth]{figures/modules.pdf}
	\caption{System Modules}
        \label{fig:modules}
\end{figure}

\begin{itemize}

\item \underline{Application module} - responsible for providing information about the relevant modules (\emph{NAT Traversal} and \emph{Signaling}) and user interface to the \emph{Client}.

\item \underline{Signaling module} - responsible for \emph{Client} and \emph{Stream} coordination.

 \item \underline{NAT Traversal module} - used by \emph{Client} and \emph{Stream} modules during the \emph{Signaling} phase which ends by establishing the connection between them.

 \item \underline{Stream module} - responsible to deliver and receive multimedia content from the \emph{Client}.

 \item \underline{Storage module} - provides two main functionalities: store the model information and media recorded. This is the single module responsible for persistent storage. It stores user and communication data as well as all the data required among user sessions. It is also use to store all the communication streams, so that they can be viewed later.

 \item \underline{Client module} - responsible for the interaction with the user.

\end{itemize}

%RP Acho que podes escrever muito mais sobre cada componente. Isto está muito curto. Fala de algumas das funções que proporcionam. Vê o que fiz com o Storage Module

%RP também podes fazer um diagrama com vários utilizadores e mostrar quais os streams que vão de e para cada um (mostrar a mistura de som, video, etc).

% Each room will be associated to multiple Media-Types, interactive and non-interactive, discrete and continuous. This media types can be structured with chains of events. Some components of the user interface may not be synchronized to all participants, for example an help or suggestion window.

\section{Implementation Proposal}
The infrastructure is composed by: web server, stream server, signaling server, database and video repository.
%RP tens da fazer o paralelo entre isto e os componentes apresentados na secção anterior. O mesmo com a imagem, que parece não estar relacionada. Podes colocar os elementos na mesma posição, ou dentro das caixas da imagem 3.1

\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{figures/infrastructure.pdf}
	\caption{System Infrastructure}
\end{figure}

In order to simplify our solution, we propose that the \emph{Application}, \emph{Signaling}, \emph{Stream} and \emph{Storage} modules be implemented within the same server application, so it could be easy to deploy as a single image. To this set of modules, we often call \emph{backend}.

Having established that the \emph{backend} modules are placed in the same machine, that helps controlling which resources the client has permission to access as those modules are seen as a private network.
%RP na mesma ``server application'' ou ``machine''? São coisa diferentes! Apresentas vantages de instalar tudo na mesma máquina mas não falas das desvantagens. Menor escalabilidade parece-me ser logo uma delas. Na figura 3.2 aparecem várias máquinas!

To qualify the above we provide public access to \ac{HTTP} server ports, maintaining the access to other components restricted through firewall rules.
%RP qualify?
%RP Imagem com firewall e estes portos? Mas isto já me parece muito detalhe de implementação.
In relation to the database there is no directly access from the outside. All the database information is accessed via our application server which validates the permissions of users on our system.

On the other hand, the access to our streaming servers is also restricted, but clients can connect to them after concluding the signaling phase. This signaling phase may or not proceed in function of the client access permissions. For example, if a user is trying to access a private conference room that he is not a member of nor has an invitation link for, the signaling server refuses to start the signaling phase and the user cannot access the streaming server.

Furthermore, we have taken into account the compatibility between the streaming server, database and the operation transformation solutions, in order choose the appropriate framework to implement our web server.
%RP esta secção devia começar aqui, a explicar a escolha das tecnologias/produtos a usar em cada componente. O que tens antes são pormenores que só faz sentido discutir no final. Tens de ir sempre do mais geral para o mais detalhado.

As a result of our study about signaling protocols we have decided to exclude \ac{XMPP} due to the development difficulties that it exposes when using multiple web pages, namely the user re-authentication performed each time a page is loaded and the single tab usage limitation. As a consequence we have not chosen \emph{Jitsi Video Brigde} for stream processing as it uses \ac{XMPP}.

Accordingly, by excluding \emph{Jitsi Video Brigde} we have decided that our solution must use \ac{KMS}. Our web server could be implemented easily with \emph{NodeJS} or \emph{Java} due to the fact \ac{KMS} provides clients for both technologies. But others could also be used as \ac{KMS} also exposes their \ac{API} via \emph{WebSockets}.

Due to the fact we are going to use \ac{KMS} as streaming server solution, we could choose \emph{NodeJS} or any \emph{Java} based web framework for implementing our web application server. One of our criteria for choosing the web framework is the ability to follow the \ac{MVC} paradigm which can help us to organize our code. Without any strong reason we have decided to implement our web server with the \emph{PlayFramework}\footnote{\url{https://www.playframework.com/}(acessed March 25, 2016)} using \emph{Java}.
%RP nem sequer por estares mais familiarizado com java ou a playframework? Ou porque querias aprender? Alguma razão há-de ter havido

By default, \emph{Kurento Repository} is implemented over \emph{MongoDB}, for convenience our storage model will also use the same database.

Importantly, for the collaborative text editor, we have chosen \emph{OT.js} due to its server and storage implementation choice independence.

For the \emph{NAT Traversal} module, an \ac{ICE} server is not required to be on our infrastructure, as a public \ac{STUN} server can be used for testing our solution. Nevertheless, we recognize that for a production environment we would need to maintain our own \ac{TURN} servers in order to ensure connectivity to all clients.

Not less important, on the client computers, both \emph{Mozilla Firefox} and \emph{Google Chrome} could be installed as web browsers. As such, both should be supported. Libraries such as \emph{jQuery}, \emph{Bootstrap}, \emph{Adapter.js}, \emph{OT.js} can be downloaded from the web server and executed on the client side using any of these two browsers.
%RP relembrar que o IE e o Safari não suportam webrtc?

\begin{table}[H]
\centering
	\caption{Application Architecture}
	\label{table:apparch}
    \begin{tabular}{cccccccc@{}m{0pt}@{}}
	\hline 
\multicolumn{8}{|c|}{\cellcolor{Gray}Application}  &\\[12pt]\cline{1-5}\cline{7-7}
\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{~~~~~ot.js~~~~~} & \multicolumn{1}{c|}{\cellcolor{Gray}~~~~~~~~~~~~~~~} & \multicolumn{1}{c|}{adapter.js} &   \multicolumn{1}{c|}{\cellcolor{Gray}~~~~~~~~~~~~~~~} &\\[12pt]\hline
\multicolumn{1}{|c|}{HTTP} & \multicolumn{2}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[12pt]\hline
\end{tabular}
\end{table}

%RP acho interessante o que queres fazer aqui mas acho que está mal conseguido. A linha de baixo, devia ter coisas ao mesmo nível, o que não aconteçe.
% Qual a razão para HTTP estar por baixo do jquery e user interface estar ao lado?
% Sugiro colocar a applicação no meio e as várias interfaces (webrtc, websocket e http) uma de cada lado (esquerda, baixo e direita), indicando o que fica do outro lado (outro cliente, servidor web, servidor stream, etc)

Table~\ref{table:apparch} presents the application architecture and the underlying technologies seen from the user's perspective. \emph{Adapter.js} and \emph{jQuery} will ensure that our application is compatible with the most popular web browsers.
\emph{Bootstrap} will be used to make the user interface more appellative and responsive. With \emph{Bootstrap} it is quite easy to develop applications that adapt to mobile devices with different screen sizes.
%RP bootstrap não aparece na figura!

The synchronization between multimedia elements will be performed through chains of \emph{JavasSript} events or by specifying the interval of time which time content must be visible. Other animations can be implemented with \ac{SVG} embedded on \ac{HTML}.
%RP Caiu do céu! Tens de explicar e ligar melhor a sequencia de texto

\section{Chapter Summary}
\label{architecture:summary}

In this chapter we have shown the modules needed in order to implement our solution.

Several technologies were taken into account for implementing each module, we have studied the pros and cons of each technology and decided the architecture for our solution.