### 2016-05-13 tese2/xacronyms.tex

\newacronym{AES}{AES}{Advanced Encryption Standard}
\newacronym{API}{API}{Application Programming Interface}
\newacronym{BOSH}{BOSH}{Bidirectional-streams Over Synchronous HTTP}
\newacronym{CBC}{CBC}{Cipher Block Chaining}
\newacronym{CPU}{CPU}{Central Processing Unit}
\newacronym{CRUD}{CRUD}{Create, Read, Update and Delete}
\newacronym{CSS}{CSS}{Cascading Style Sheets}
\newacronym{DOM}{DOM}{Document Object Model}
\newacronym{DTLS}{DTLS}{Datagram Transport Layer Security}
\newacronym{DVD}{DVD}{Digital Versatile Discs}
\newacronym{FMS}{FMS}{Flash Media Server}
\newacronym{HTML}{HTML}{HyperText Markup Language}
\newacronym{HTTP}{HTTP}{Hypertext Transfer Protocol}
\newacronym{ICE}{ICE}{Interactive Connectivity Establishment}
\newacronym{IM}{IM}{Instant Messenger}
\newacronym{IoT}{IoT}{Internet of Things}
\newacronym{IPv4}{IPv4}{Internet Protocol Version 4}
\newacronym{IPv6}{IPv6}{Internet Protocol Version 6}
\newacronym{IP}{IP}{Internet Protocol}
\newacronym{JSON}{JSON}{JavaScript Object Notation}
\newacronym{JVM}{JVM}{Java virtual machine}
\newacronym{KMS}{KMS}{Kurento Media Server}
\newacronym{MCU}{MCU}{Multipoint Control Unit}
\newacronym{MPEG}{MPEG}{Moving Picture Experts Group}
\newacronym{MVC}{MVC}{Model-View-Controller}
\newacronym{NAT}{NAT}{Network Address Translation}
\newacronym{OT}{OT}{Operational transformation}
\newacronym{P2P}{P2P}{Peer-to-peer}
\newacronym{PSTN}{PSTN}{Public Switched Telephone Network}
\newacronym{RAM}{RAM}{Random-access memory}
\newacronym{REST}{REST}{Representational State Transfer}
\newacronym{RFC}{RFC}{Request For Comments}
\newacronym{RTCP}{RTCP}{Real Time Control Protocol}
\newacronym{RTSP}{RTSP}{Real Time Streaming Protocol}
\newacronym{RTMFP}{RTMFP}{Real-Time Media Flow Protocol}
\newacronym{RTMP}{RTMP}{Real Time Messaging Protocol}
\newacronym{RTP}{RTP}{Real-time Transport Protocol}
\newacronym{SAMI}{SAMI}{Synchronized Accessible Media Interchange}
\newacronym{SCTP}{SCTP}{Stream Control Transmission Protocol}
\newacronym{SDP}{SDP}{Session Description Protocol}
\newacronym{SigOfly}{SigOfly}{Signaling-On-the-fly}
\newacronym{SIMPLE}{SIMPLE}{SIP for Instant Messaging and Presence Leveraging Extensions}
\newacronym{SIP}{SIP}{Session Initiation Protocol}
\newacronym{SMIL}{SMIL}{Synchronized Multimedia Integration Language}
\newacronym{SMS}{SMS}{Short Message Service}
\newacronym{SRTCP}{SRTCP}{Secure RTCP}
\newacronym{SRTP}{SRTP}{Secure Real-time Transport Protocol}
\newacronym{SRT}{SRT}{SubRip Text}
\newacronym{SSL}{SSL}{Secure Sockets Layer}
\newacronym{STUN}{STUN}{Session Traversal Utilities for NAT}
\newacronym{SVG}{SVG}{Scalable Vector Graphics}
\newacronym{TCP}{TCP}{Transmission Control Protocol}
\newacronym{TLS}{TLS}{Transport Layer Security}
\newacronym{TMN}{TMN}{This Means Nothing}
\newacronym{TURN}{TURN}{Traversal Using Relays around NAT}
\newacronym{UDP}{UDP}{User Datagram Protocol}
\newacronym{URL}{URL}{Uniform Resource Locator}
\newacronym{VoIP}{VoIP}{Voice Over IP}
\newacronym{WebRTC}{WebRTC}{Web Real-Time Communication}
\newacronym{WebVTT}{WebVTT}{Web Video Text Tracks}
\newacronym{XAML}{XAML}{eXtensible Application Markup Language}
\newacronym{XEP}{XEP}{XMPP Extensions}
\newacronym{XHTML}{XHTML}{eXtensible Hypertext Markup Language}
\newacronym{XML}{XML}{eXtensible Markup Language}
\newacronym{XMPP}{XMPP}{Extensible Messaging and Presence Protocol}