### 2016-05-22 tese2/implementation.tex

\chapter{Implementation}
\label{chapter:implementation}

%\section{Implementation Options}
%Neste secção devem apresentar as opções de implementação que tinham ao vosso
%dispor, avaliá-las e justificar a escolha que realizaram. Isto pode englobar:
%\begin{itemize}
%    \item Simuladores
%    \item Linguagens e ambientes de programação
%    \item Sistemas operativos
%    \item Hardware
%\end{itemize}
%Fica sempre bem na avaliação colocar uma tabela com as características
%pretendidas e as que são satisfeitas pelas várias opções (tipo catálogo com as
%características dos automóveis). A escolha deve surgir naturalmente, com base na
%opção que tem mais cruzes\ldots

%\section{Architecture}
%Nesta secção devem explicar como implementaram a vossa solução, apresentando
%as simplificações que efectuaram, face ao modelo inicialmente previsto. As
%simplificações devem ser devidamente justificadas. Se for possível, devem indicar
%que estas não põem em causa as contribuições da tese.
%Podem ainda descrever os principais problemas que tiveram e a forma como os
%abordaram e resolveram.

%Se estiverem a usar um simulador devem:

%\begin{itemize}
%    \item explicar o funcionamento do simulador
%    \item explicar as alterações e modelos que desenvolveram no simulador e que permitem validar a vossa ideia
%\end{itemize}

%Se estiveram a desenvolver SW, sem simulador devem:
%\begin{itemize}
%    \item explicar os módulos, interfaces, estruturas de dados, etc\ldots
%\end{itemize}

%Sempre que possível, ilustrem a arquitectura com figuras que demonstrem a
%evolução face à arquitectura da secção anterior. Isto é, usem as figuras anteriores
%e façam as modificações necessárias à obtenção da arquitectura do protótipo.
%\ldots

In this chapter we are going to present our implementation choices, the difficulties that we have faced and the strategies available to overcome them.

To this end, we are going to present and explain our database model, how we have improved it and how this model can help us reach our goals.

Subsequently, we are going to describe our signaling protocol which will make it possible to use the \ac{WebRTC}'s functionalities in order to implement our system's streaming features.

With the model and signaling protocol defined we are prepared to show the multiple approaches we have followed in order to implement stream recording, their drawbacks and our choices.

Afterwards we are going to describe how we show hyper-content to users, namely the multiple ways to create hyper-content and the algorithm behind the user interface synchronization, the security flaws of our content displaying mechanism and a possible solution to overcome vulnerabilities.

Not less important we also are going to describe how we have implemented our application timeline and the functionalities above it such as time manipulation and annotations management. Among other different functionalities we also are going to describe how we perform stream composition, synchronization of our collaborative text editor and implementation of our instant messenger.

Lastly, we are going to describe the deployment of our solution in respect to the hardware that we have used and the software installed onto it.

\input{13_implementation_model.tex}

\section{XMPP experiments}
Although we have used our own signaling protocol on our implementation, the first steps we made during the development of our prototype were using \ac{XMPP}. We discovered the drawbacks of this approach in practice and described our learnings in the state of the art. We have found and fixed a bug present on \emph{strophe.js} library that silenced an error raised by duplicated user registrations.

As a consequence of not using \ac{XMPP} for the signaling protocol, we have defined our own protocol based on \emph{WebSockets} for performing the stream between clients and \ac{KMS}. From the \ac{KMS} functionalities, our solution uses the recording feature, composite endpoints that mix multiple audio and video streams into a single one and \ac{QR} code detection for creating hyper content.

\input{14_implementation_signaling.tex}
\input{15_implementation_recording.tex}
\input{16_implementation_content.tex}
\input{18_implementation_summary.tex}