### 2016-06-09 tese2/frontCover.tex

\thispagestyle{empty}

% IST Logo
% parameters: bb=llx lly urx ury (bounding box), width=h_length, height=v_length, angle=angle, scale=factor, clip=true/false, draft=true/false. 
\newgeometry{left=2cm,top=2.5cm, right=2.5cm, bottom=0cm}
\includegraphics[width=5.0cm]{Logo.png}

\begin{center}
%
% Figure (Image or plot)
\vspace{0.3cm}
% TODO - Use of an image is optional
\hspace{-0.5cm}\includegraphics[height=50mm]{Figures/logo.png}
% TODO - If you do not use an image uncomment the next line
% \vspace{5.0cm}

% Title, author and degree
\vspace{0.8cm}
{\FontLb Hyper-linked Communications: WebRTC enabled asynchronous collaboration} \\
%\vspace{0.2cm}
%{\FontMn Subtítulo (facultativo)} \\
%\vspace{1.9cm}
\vspace{2.1cm}
{\FontMb Henrique Duarte Lopes Rocha} \\
\vspace{1.4cm}
{\FontLn Dissertation submitted to obtain the Master Degree in} \\
\vspace{0.3cm}
{\FontLb Communications Networks Engineering} \\
\vspace{0.5cm}
{\FontSn %
Supervisors:        Prof. Ricardo Jorge Feliciano Lopes Pereira \\
        Eng.  Paulo Guilherme Gonçalves Chainho \\
}
\vspace{1.9cm}
{\FontMb Examination Committee} \\
\vspace{0.3cm}
{\FontSn %
Chairperson: 	   Prof. Paulo Jorge Pires Ferreira \\
Supervisor:        Prof. Ricardo Jorge Feliciano Lopes Pereira \\
Members of the Committee: Prof. João Pedro Barreto
}
\vspace{1.5cm}

{\FontMb May 2016} \\
%
\end{center}

\cleardoublepage

\restoregeometry